---
title: "Lab 8 - Linear Regression in R"
author: "Tara Jagadeesh"
date: "November 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

salmon <- read_csv("salmon_imports.csv")
```

##1. Exploratory scatterplot
```{r}
salmon_exploratory <- ggplot(salmon, aes(x=year, y = salmon_fresh))+
  geom_point()
salmon_exploratory

#Looks as though there could be a linear relationship
#There are, however, a few years which might not fit this trend
```

##2. Linear regression
```{r}
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model
```

Imports(tons) = 7550(Year) - 14982940   

The slope (7550) means that every year the expected salmon imports increase by 7550 tons. 

The y intercept means that, at year 0, -14982940 tons of imports are expected. This is concerning because import values cannot be negative. When extrapolating this model, this means that at a certain year in the past salmon imports stop declining and reach zero. Thus, this model is only valid for a specified time frame.     

##3. Model diagnostics
```{r}
par(mfrow = c(2,2))
plot(salmon_model)
```
The residuals do appear normally distributed, but there are a couple outliers. The residuals also appear potentially heteroscedastic. 

##3. Assessing model fit
```{r}
summary(salmon_model)
```

p < 0.001
R2 = 0.9478

1. Year significantly predicts salmon imports (p < 0.001)
2. R2 = 0.9478. This means that 94% of the variance in salmon imports is predicted by year. 
3. 6% of the variance in salmon imports is NOT explained by year. 

##4. Final graph
```{r}
salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
geom_point() +
geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
theme_bw() +
scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")
salmon_final_graph
```

##5. Making predictions
```{r}
new_years <- data.frame(year = c(2022, 2024, 2026, 2028)) #Make new dataframe to predict

future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")
future_predict

predictions <- data.frame(new_years, future_predict) #Bind the two tables together
predictions
```
##6. Pearson's r
```{r}
pearsons_r <- cor.test(salmon$year, salmon$salmon_fresh)
pearsons_r
```

pearson's r = 0.974

There is a strong positive correlation between year and tons of salmon imported (Pearson's r = 0.974)

##7. Conclusion statements

Year significantly predicts tons of salmon imported (b = 7550, t(26) = 21.7, p < 0.001) with a strong positive correlation between the two (Pearson’s r = 0.97). The overall model (imports = 7550(year) - 14982940) explains a significant amount of variance in salmon imports (F(1,26) = 471.9, p < 0.001, R^2^ = 0.94).




